Performance Models for Data Transfers: A Case Study with Molecular Chemistry Kernels

### 2019-03-27 Article/article.tex
Experimental Results and Plots

	We consider a machine, Cascade, available at PNNL, for our experiments. We obtain traces by running  two molecular chemistry applications,  double precision version of Hartree-Fock (HF) and Coupled Cluster Single Double (CCSD) of NWChem package on 10 nodes of this machine. Each node is composed of 16 Intel Xeon E5-2670 cores, hence we use 160 processes for the execution. NWChem takes advantages of a Partitioned Global Address Space Programming Model, Global Arrays (GA), to use shared-memory programming APIs  on distributed memory computers.  Our data transfer model is quite simple and  we consider that all data transfers between each process's local memory and GA memory take the same route. Modeling of different routes of data transfers for the same source-destination pair, bandwidth sharing for different source-destination pairs and network congestion is more challenging and part of our future work. Our model is simple yet it provides insight to the application developers (or runtime system) that  in what-order all data transfer  are issued for the same source-destination pair to maximise communication-computation overlap. Our model is easily adaptable to any source-destination where there is one fixed route between source and destination (such as between CPU and GPU, one copy engine to transfer data from CPU (resp. GPU) to GPU (resp. CPU) ).
	\todo[inline]{More about tiling}
	Both application mainly perform two types of computations, tensor transpose and tensor contractions.  HF expects to specify a tile size and we set it to 100, so that each core can be efficiently utilize.  CCSD automatically determines tile sizes at different program points based on the input (molecule?).  Hence, HF operates on almost homogeneous tiles while CCSD on heterogeneous tiles.

	\todo[inline]{Scheduled based on Gilmore-Gomory algorithm and its performance }
	\todo[inline]{Skumar: HF performance plots}
	\begin{figure}[htb]
		\includegraphics[scale=0.25]{../HFPlots/ratio_to_optimal.pdf}
		\caption{HF Performance.}
		\label{fig:staticOrderDynamicCorrectionsExample}
	\end{figure}
	\begin{figure}[htb]
	\includegraphics[scale=0.25]{../CCSDPlots/best_ccsd.pdf}
	\caption{HF best Performance.}
	\label{fig:staticOrderDynamicCorrectionsExample}
\end{figure}
	\todo[inline]{Skumar: CCSD performance plots}
		\begin{figure}[htb]
		\includegraphics[scale=0.25]{../CCSDPlots/ratio_to_optimal_ccsd.pdf}
		\caption{CCSD Performance.}
		\label{fig:staticOrderDynamicCorrectionsExample}
	\end{figure}
	\begin{figure}[htb]
		\includegraphics[scale=0.25]{../CCSDPlots/best_ccsd.pdf}
		\caption{CCSD best Performance.}
		\label{fig:staticOrderDynamicCorrectionsExample}
	\end{figure}
	\todo[inline]{Skumar: Effect of Random memory with HF or CCSD}

>>>>

	We consider a machine called Cascade, available at PNNL, for our experiments. We obtain traces by running  two molecular chemistry applications,  double precision version of Hartree-Fock (HF) and Coupled Cluster Single Double (CCSD) of NWChem package with 4 Uracil molecules input on 10 nodes of this machine. Each node is composed of 16 Intel Xeon E5-2670 cores, hence we use 160 processes for the execution. Each process executes 300-800 tasks.  NWChem takes advantages of a Partitioned Global Address Space Programming Model, Global Arrays (GA), to use shared-memory programming APIs  on distributed memory computers.  Our data transfer model is quite simple and  we consider that all data transfers between each process's local memory and GA memory take the same route. Modeling of different routes of data transfers for the same source-destination pair, bandwidth sharing for different source-destination pairs and network congestion is more challenging and part of our future work. Our model is simple yet it provides insight to the application developers (or runtime system) that  in what-order all data transfer  are issued for the same source-destination pair to maximise communication-computation overlap. Our model is easily adaptable to any source-destination where there is one fixed route between source and destination (such as between CPU and GPU, one copy engine to transfer data from CPU (resp. GPU) to GPU (resp. CPU) ).
	\todo[inline]{More about tiling}
	Both application mainly perform two types of computations, tensor transpose and tensor contractions.  HF expects to specify a tile size and we set it to 100, so that each core can be efficiently utilize.  CCSD automatically determines tile sizes at different program points based on the input (molecule?).  Hence, HF operates on almost homogeneous tiles while CCSD on heterogeneous tiles.

%%	\todo[inline]{Scheduled based on Gilmore-Gomory algorithm and its performance }
	We also consider an algorithm, proposed by Gilmore and Gomory, to obtain the minimal cost sequence for a set of jobs. In this algorithm, each job has a start and end state and a cost is associated to change the state. In our context, this cost can be seen as non-overlap time of computation for two adjacent jobs. Here is the main idea of this algorithm. Initially partial sequence of jobs is represented by a graph such that their overlap is maximum. Subsequently edges are added to this graph which minimized the total non-overlap cost  and connects two components. When all nodes of this graph is connected then an edge interchange mechanism is taken into account to determine the sequence of jobs, which ensures that the sequence has minimal cost.

	We refer ~\cite{Gilmore-Gomory:1964} for the detailed algorithm and ~\cite{gitworkrepo} for our implementation. This algorithm does not take memory capacity into consideration and only provides the sequence of processing. We applied the same sequence with different memory capacity restrictions for our experiments and call this heuristic as \textit{Gilmore-Gomory}.

	\todo[inline]{Skumar: HF performance plots}
	\begin{figure}[htb]
		\includegraphics[scale=0.25]{../HFPlots/ratio_to_optimal.pdf}
		\caption{HF Performance.}
		\label{fig:staticOrderDynamicCorrectionsExample}
	\end{figure}
	\begin{figure}[htb]
	\includegraphics[scale=0.25]{../CCSDPlots/best_ccsd.pdf}
	\caption{HF best Performance.}
	\label{fig:staticOrderDynamicCorrectionsExample}
\end{figure}
	\todo[inline]{Skumar: CCSD performance plots}
		\begin{figure}[htb]
		\includegraphics[scale=0.25]{../CCSDPlots/ratio_to_optimal_ccsd.pdf}
		\caption{CCSD Performance.}
		\label{fig:staticOrderDynamicCorrectionsExample}
	\end{figure}
	\begin{figure}[htb]
		\includegraphics[scale=0.25]{../CCSDPlots/best_ccsd.pdf}
		\caption{CCSD best Performance.}
		\label{fig:staticOrderDynamicCorrectionsExample}
	\end{figure}
	\todo[inline]{Skumar: Effect of Random memory with HF or CCSD}
